---
title: "Geopolitics"
format:
  dashboard:
    theme: minty
    logo: logo.png
    nav-buttons:
      - linkedin
      - twitter
      - github
    orientation: columns
embed-resources: true
execute: 
  warning: false
  code-fold: true
server: shiny
---


```{r}
#| context: setup
#| warning: false
library(tidyverse)
library(ggplot2)
library(ggridges)
library(lmtest)
library(gt)
library(countrycode)
library(shiny)

democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-05/democracy_data.csv')
```

```{r}
#| context: data
#| results: 'asis'

democracy_data <- democracy_data |> 
  mutate(
    continent = case_when(
      
      country_code %in% c(
        "DZA","AGO","BEN","BWA","BFA","BDI","CMR","CPV","CAF","TCD","COM","COG","COD",
        "CIV","DJI","EGY","GNQ","ERI","SWZ","ETH","GAB","GMB","GHA","GIN","GNB","KEN",
        "LSO","LBR","LBY","MDG","MWI","MLI","MRT","MUS","MAR","MOZ","NAM","NER","NGA",
        "RWA","STP","SEN","SYC","SLE","SOM","ZAF","SSD","SDN","TZA","TGO","TUN","UGA",
        "ZMB","ZWE","ZAR"
      ) ~ "Africa",

      # North America (incl. Caribbean)
      country_code %in% c(
        "CAN","USA","MEX","BHS","BRB","CUB","DMA","DOM","GRD","HTI","JAM",
        "KNA","LCA","VCT","TTO","ATG"
      ) ~ "North America",

      # Central America
      country_code %in% c(
        "BLZ","CRI","SLV","GTM","HND","NIC","PAN"
      ) ~ "Central America",

      # South America
      country_code %in% c(
        "ARG","BOL","BRA","CHL","COL","ECU","GUY","PRY","PER","SUR","URY","VEN"
      ) ~ "South America",

      country_code %in% c(
        "AFG","ARM","AZE","BHR","BGD","BTN","BRN","KHM","CHN","CYP","GEO","IND","IDN",
        "IRN","IRQ","ISR","JPN","JOR","KAZ","KWT","KGZ","LAO","LBN","MYS","MDV","MNG",
        "MMR","NPL","PRK","OMN","PAK","PSE","PHL","QAT","SAU","SGP","KOR","LKA","SYR",
        "TWN","TJK","THA","TLS","TUR","TKM","ARE","UZB","VNM","YEM"
      ) ~ "Asia",

      country_code %in% c(
        "ALB","AND","AUT","BLR","BEL","BIH","BGR","HRV","CYP","CZE","DNK","EST","FIN",
        "FRA","DEU","GRC","HUN","ISL","IRL","ITA","LVA","LIE","LTU","LUX","MLT","MDA",
        "MCO","MNE","NLD","MKD","NOR","POL","PRT","ROU","RUS","SMR","SRB","SVK","SVN",
        "ESP","SWE","CHE","UKR","GBR","VAT","GER","ROM"
      ) ~ "Europe",

      country_code %in% c(
        "AUS","FJI","KIR","MHL","FSM","NRU","NZL","PLW","PNG","WSM","SLB","TON","TUV",
        "VUT","NUR"
      ) ~ "Oceania",

      TRUE ~ NA_character_
    )
  ) |> 
  filter(!is.na(continent))
```

```{r}
#| context: data
#| label: population-table

library(rvest)

populations <- read_html("https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population")

pop_data <- populations |>
  html_element("table.wikitable") |>
  html_table(fill = TRUE) |>
  select(
    Country = 1, 
    Population = 2
    ) |>
  mutate(
    Country    = trimws(Country),
    Population = as.numeric(gsub("[^0-9]", "", Population))  # strip anything that isn't a digit
  ) |>
  filter(!is.na(Population))

```

```{r}
#| context: data
#| label: country-data

pop_data <- pop_data |>
  mutate(
    iso3c = countrycode(Country, 
                        origin = "country.name", 
                        destination = "iso3c"
                        ))

democracy_data_2020 <- democracy_data |>
  filter(year == 2020) |>
  mutate(
    iso3c = countrycode(country_name, 
                        origin = "country.name", 
                        destination = "iso3c"
                        ))

# Now join based on iso3c country code
country_data <- pop_data |>
  inner_join(democracy_data_2020, 
             by = "iso3c")

```

```{r}
#| context: data
#| label: country-api

library(glue)
library(jsonlite)
library(httr)

# Obtain 2020 GDP in USD for inputted country
get_gdp_2020 <- function(country, key = my_key)
  {
  
  # Store url to access API with country and key queries
  url <- glue("https://api.api-ninjas.com/v1/gdp?year=2020;country={country};X-Api-Key={key}")
  
  # Temporarily store JSON data safely
  my_data <- safely(fromJSON)(url)
  
  # Check if JSON was successfully obtained
  if (is.null(my_data$result) |
      length(my_data$result) == 0) {
    return(NA)
  } else {
    return(my_data$result$gdp_nominal)
  }
}

```

```{r}
#| context: data
#| echo: false
my_key <- "Co1aNrJSrFjlk3zmHb2jrVqKvyfcBq4wJ5TP3JMo"
```


```{r}
#| context: data
#| eval: false
# We used a personal API key stored as "my_key"

# Get GDP for all countries
gdps <- map_dbl(.x = country_data$iso3c, .f = get_gdp_2020)

# Store GDP with country code as a tibble
country_gdps <- tibble(iso3c = country_data$iso3c, GDP = gdps)

# Save tibble as a csv
write.csv(country_gdps, file = "country_gdps.csv")
```

```{r}
#| context: data
# Read in csv GDP data
gdp_data <- read.csv("country_gdps.csv")
gdp_data <- gdp_data |>
  select(-X)

# Combine GDP with aggregate 2020 data
country_meta_data <- country_data |>
  inner_join(gdp_data, 
             by = "iso3c", 
             relationship = "many-to-many")

```

```{r}
#| context: data
country_meta_data <- country_meta_data |>
  filter(!is.na(regime_category_index)) |> 
  mutate(gov_type = case_when(
    is_communist == TRUE ~ "Communist",
    is_democracy == TRUE & is_monarchy == TRUE ~ "Democratic Monarchy",
    is_democracy == TRUE & is_monarchy == FALSE ~ "Pure Democracy",
    is_democracy == FALSE & is_monarchy == TRUE ~ "Pure Monarchy",
    TRUE ~ "Other"),
    gov_type = factor(gov_type, levels = c("Pure Democracy",
                                           "Pure Monarchy",
                                           "Democratic Monarchy",
                                           "Communist",
                                           "Other"
                                           )),
    regime_category = case_when(
    regime_category_index == 0 ~ "Parliamentary democracy",
    regime_category_index == 1 ~ "Mixed democratic", 
    regime_category_index == 2 ~ "Presidential democracy",
    regime_category_index == 3 ~ "Civilian dictatorship", 
    regime_category_index == 4 ~ "Military dictatorship",
    regime_category_index == 5 ~ "Royal dictatorship")
  )

```

```{r}
#| context: data
plot_gdp_by_category <- function(cat_var, data = country_meta_data) {
  data |>
    filter(!is.na({{ cat_var }})) |> 
    ggplot(aes(y = {{ cat_var }}, x = GDP, fill = {{ cat_var }})) + 
    geom_density_ridges(alpha = 0.5) + 
    scale_fill_viridis_d() + 
    scale_x_log10(labels = scales::comma) +
    labs(
      x     = "GDP (log scale)",
      y     = "", 
      title = paste("Country GDP in USD (in Billions) by Government Type"), 
      subtitle = paste("Comparing the GDP of Countries by Government Type")
    ) +
    theme_bw() + 
    theme(legend.position = "none", 
          panel.grid.minor = element_blank(), 
          text = element_text(family = "Helvetica", size = 12))
  
}
```

```{r}
#| context: data
summarize_gdp_by_category <- function(cat_var, data = country_meta_data) {
  
  data |>
    filter(!is.na({{ cat_var }})
    ) |>
    group_by({{ cat_var }}) |>
    summarize(n = n(),
              mean_GDP = mean(GDP, na.rm = TRUE),
              med_GDP = median(GDP, na.rm = TRUE),
              sd_GDP = sd(GDP, na.rm = TRUE),
              se_GDP = sd_GDP / sqrt(n),
              min_gdp = min(GDP, na.rm = TRUE),
              max_gdp = max(GDP, na.rm = TRUE),
              .groups    = "drop"
    ) |>
    arrange(desc(mean_GDP)) |>
    gt() |>
    fmt_currency(columns = 3:8,
                 decimals = 2,
                 currency = "USD",
                 placement = "left",
                 pattern = "{x}B"
    ) |>
    tab_spanner(label = "Category Breakdown",
                columns = 1:2
    ) |>
    tab_spanner(label = "Midpoints",
                columns = 3:4
    ) |>
    tab_spanner(label = "Variation",
                columns = 5:6
    ) |>
    tab_spanner(label = "Range",
                columns = 7:8
    ) |>
    cols_label(gov_type = "Government Type",
               n = "Count",
               mean_GDP = "Mean",
               med_GDP = "Median",
               sd_GDP = "Standard Deviation",
               se_GDP = "Standard Error",
               min_gdp = "Minimum",
               max_gdp = "Maximum"
    ) |>
    data_color(method = "numeric",
               palette = "Blues"
    ) |>
    tab_header(title = paste("Table of GDP Statistics by Government Type"),
               subtitle = "(in USD in Billions)"
    ) |>
    tab_style(style = cell_text(weight = "bold"),
              locations = list(cells_title(groups = "title"), 
                               cells_title(groups = "subtitle"), 
                               cells_column_labels())
    ) |>
    tab_footnote(md("*Note 1*: Standard error was calculated as the standard deviation divided by the square root of the sample size.")) |>
    tab_footnote(md("*Note 2*: Darker blues represent greater values."))
  }

```

# Main Research Question

## Column 1
:::{.card title = "Are certain political ideologies more popular among higher/lower GDP countries?"}

::: {.toolbar}


```{r}
checkboxGroupInput(inputId = "GovTypes",
              label = "Select Government Types to Compare:",
              choices = unique(country_meta_data$gov_type),
              selected = unique(country_meta_data$gov_type))

```

:::

::: {.panel-tabset}

### Ridgeline Plot

```{r}
plotOutput("gov_type_plot")
```

### Summary Table

```{r}
gt_output("gov_type_table")
```

:::

:::

## Column 2

:::{.card title = "Leaflet Plot - Coloring each country by Government Type"}

Magnitude of GDP represented by the size of the orange circle. GDP is scaled to radius = $\frac{\sqrt{GDP}}{5}$

```{r}
#| cache: true
library(sf)
library(leaflet)
library(rnaturalearth)
library(rnaturalearthdata)

global_map <- ne_countries(scale = "medium", returnclass = "sf")

center_coords <- global_map |>
  st_centroid() |>
  st_coordinates() |>
  as.tibble()

country_centers <- tibble(iso_a3 = global_map$iso_a3,
                          lon = center_coords$X,
                          lat = center_coords$Y
) |>
  select(iso_a3, lon, lat) |>
  filter(!is.na(lon),
         !is.na(lat))

world_map <- global_map |>
  left_join(country_centers,
            by = "iso_a3")



global_gdps <- country_meta_data |>
  select(Country, gov_type, GDP, iso3c)

GDP_with_Coords <- global_gdps |>
  left_join(world_map,
            by = c("iso3c" = "iso_a3"))

GDP_with_Coords <- st_as_sf(GDP_with_Coords)

gov_cols <- colorFactor(
  palette = "viridis",
  domain = GDP_with_Coords$gov_type
)

GDP_with_Coords |>
  leaflet() |>
  addTiles() |>
  setView(lat = 0, lng = 0, zoom = 1.4) |>
  addPolygons(color = "black",
              weight = 1,
              opacity = 1,
              fillColor = ~gov_cols(gov_type),
              fillOpacity = 1,
  ) |>
  addLegend(pal = gov_cols,
            values = ~factor(gov_type,
                             levels = c("Communist",
                                        "Pure Monarchy",
                                        "Democratic Monarchy",
                                        "Pure Democracy",
                                        "Other")
            ),
            title = "Government Type"
  ) |>
  addCircleMarkers(radius =  ~sqrt(GDP) / 5,
                   lng = ~lon,
                   lat = ~lat,
                   weight = 2,
                   color = "#EF7924",
                   fillColor = "#EF7924",
                   opacity = 1,
                   fillOpacity = 0.6,
                   label = ~glue("{name}: GDP = ${GDP}B"),
                   labelOptions = labelOptions(textsize = "15px")
  )

```

:::


# Focusing on China

## Column 1

::: {.card title = "Why China?"}

Looking at the GDP data, we see that the 'Communist' Government Type has a very high average along with a high standard deviation. Looking deeper into the individual 2020 GDP country data, we noticed that China was a huge outlier and decided to take a look into the GDP of China from 1981-2020.

```{r}
#| context: data
country_meta_data |> 
  filter(gov_type == "Communist") |> 
  select(country_name, GDP) |> 
  slice_max(order_by = GDP, n = 5) |> 
  gt() |> 
  tab_header(
    title = "GDP's of Communist Countries"
  ) |> 
  cols_label(
    country_name = "Country"
  ) |>
  fmt_currency(columns = GDP) |>
  tab_options(
    table.width = pct(90)
  )


# Obtain 2020 GDP in USD for inputted country
get_gdp_china <- function(year, key = my_key)
  {
  
  # Store url to access API with country and key queries
  url <- glue("https://api.api-ninjas.com/v1/gdp?country=China&year={year}&X-Api-Key={key}")
  

  
  # Temporarily store JSON data safely
  my_data <- safely(fromJSON)(url)
  
  # Check if JSON was successfully obtained
  if (is.null(my_data$result) |
      length(my_data$result) == 0) {
    return(NA)
  } else {
    return(my_data$result$gdp_nominal)
  }
}
```

```{r}
#| context: data
#| eval: false

# Get GDP for China
china_gdps <- map_dbl(.x = 1981:2020, .f = get_gdp_china)

# Store GDP with years as a tibble
china_gdps <- tibble(years = 1981:2020, GDP = china_gdps)

# Save tibble as a csv
 write.csv(china_gdps, file = "china_gdps.csv")
```

```{r}
#| context: data
# Read in csv GDP data
china_gdps <- read.csv("china_gdps.csv")

china_gdps <- china_gdps |>
  select(-X)

```

:::

## Column 2

:::{.card title = "Animated Line Chart"}

```{r}
#| cache: true
library(gganimate)
library(gifski)

p <- china_gdps |> 
  ggplot(aes(x = years, y = GDP)) + 
  geom_area(stat = "identity",
            fill = "#EF7924") +
  geom_line(color = "black") + 
  geom_point() +
  theme(legend.position = "none") +
  theme_bw() +
  labs(
    title = "GDP of China in USD from 1981-2020",
    x = "Year",
    y = "GDP (Billions)") +
  transition_reveal(years)

china_anim <- animate(p, nframes = 100, fps = 10, end_pause = 30, renderer = gifski_renderer(tempfile(fileext = ".gif")))
anim_save("china_gdps_over_time.gif", china_anim)

china_anim
```

```{r}
#| context: server

filtered_gov_types <- reactive({
  req(input$GovTypes)
  filt_countries <- country_meta_data |>
    filter(gov_type %in% input$GovTypes)
  filt_countries
})

output$gov_type_plot <- renderPlot({
  filtered_gov_types() |>
    ggplot(aes(y = gov_type,
               x = GDP,
               fill = gov_type)
           ) + 
    geom_density_ridges(alpha = 0.5) + 
    scale_fill_viridis_d() + 
    scale_x_log10(labels = scales::comma) +
    labs(
      x     = "GDP (log scale)",
      y     = "", 
      title = "Country GDP in USD (in Billions) by Government Type", 
      subtitle = "Comparing the GDP of Countries by Government Type"
      ) +
    theme_bw() + 
    theme(legend.position = "none", 
          panel.grid.minor = element_blank(), 
          text = element_text(family = "Helvetica", size = 12))
})



output$gov_type_table <- render_gt({
  filtered_gov_types() |>
    group_by(gov_type) |>
    summarize(n = n(),
              mean_GDP = mean(GDP, na.rm = TRUE),
              med_GDP = median(GDP, na.rm = TRUE),
              sd_GDP = sd(GDP, na.rm = TRUE),
              se_GDP = sd_GDP / sqrt(n),
              min_gdp = min(GDP, na.rm = TRUE),
              max_gdp = max(GDP, na.rm = TRUE),
              .groups    = "drop"
    ) |>
    arrange(desc(mean_GDP)) |>
    gt() |>
    fmt_currency(columns = 3:8,
                 decimals = 2,
                 currency = "USD",
                 placement = "left",
                 pattern = "{x}B"
    ) |>
    tab_spanner(label = "Category Breakdown",
                columns = 1:2
    ) |>
    tab_spanner(label = "Midpoints",
                columns = 3:4
    ) |>
    tab_spanner(label = "Variation",
                columns = 5:6
    ) |>
    tab_spanner(label = "Range",
                columns = 7:8
    ) |>
    cols_label(gov_type = "Government Type",
               n = "Count",
               mean_GDP = "Mean",
               med_GDP = "Median",
               sd_GDP = "Standard Deviation",
               se_GDP = "Standard Error",
               min_gdp = "Minimum",
               max_gdp = "Maximum"
    ) |>
    data_color(method = "numeric",
               palette = "Blues"
    ) |>
    tab_header(title = paste("Table of GDP Statistics by Government Type"),
               subtitle = "(in USD in Billions)"
    ) |>
    tab_style(style = cell_text(weight = "bold"),
              locations = list(cells_title(groups = "title"), 
                               cells_title(groups = "subtitle"), 
                               cells_column_labels())
    ) |>
    tab_footnote(md("*Note 1*: Standard error was calculated as the standard deviation divided by the square root of the sample size.")) |>
    tab_footnote(md("*Note 2*: Darker blues represent greater values."))
})

```

